---
title: "Read & Format SU York data"
author: "Kim Engie"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(vroom)
```

### Read in csvs, and make a data dictionary of release types

```{r}
basedir = "/Users/kimengie/Documents/Data/York/jail"
cat(file=stderr(), "\nreading data...\n")

inmates <- vroom(file.path(basedir, "yorkopenlattice_withbookingnumbers.csv"),
                  col_names = c("first", "last", "status", "statusdate", "id", "pouch"))

bookings <- vroom(file.path(basedir, "YCPInmateHistory2017.csv"))

releasetype = tribble(~ReleaseType, ~Description,
                      "BA","BY BAIL",
                        "CO", "BY COURT'S ORDER (COURT OF RECORD AND MINOR JUDICIARY)",
                        "DC", "SENTENCED TO THE DEPARTMENT OF CORRECTIONS",
                        "DS", "BY DEATH",
                        "EM", "BY EXPIRATION OF MAXIMUM",
                        "FC", "BY PAYMENT OF FINES AND COSTS",
                        "HA", "BY HOLDING AUTHORITY",                                                
                        "OT", "OTHER",
                        "PC", "PAROLED BY THE COURT OF RECORD",
                        "PS", "PAROLED BY STATE BOARD OF PROBATION/PAROLE",
                        "PT", "PERMANENT TRANSFER TO OTHER INSTITUTION",
                        "VA",  "BY VARIOUS FEDERAL/STATE/POLICE AUTHORITIES)")
```

More csvs - from sftp folder
```{r}
# File list 
allfiles <- list.files(path=basedir, pattern="yorkcounty_\\d+.csv", full.names=TRUE, recursive=FALSE)

for (i in allfiles){
  name = str_split(i, "/")  %>% tail() %>% pluck(1,9) %>% str_remove(".csv")
  assign(name, vroom(i, col_types = cols(.default = "c")))
}

# test = "/Users/kimengie/Documents/Data/York/jail//yorkcounty_202201031600.csv"
# str_split(test, "/") %>% tail() %>% pluck(1,9) %>% str_remove(".csv") %>% str_replace("\"", "")
```

Stack the recent csvs as they're all the same format, which is the same format as the bookings table
```{r}
sftp_bookings <- rbind(yorkcounty_202111021124, 
                     yorkcounty_202111230945,
                     yorkcounty_202111231000, 
                     yorkcounty_202111231015,
                     yorkcounty_202111231030,
                     yorkcounty_202111231045,
                     yorkcounty_202111231100,
                     	yorkcounty_202112101115,
                     yorkcounty_202112101130,
                     yorkcounty_202112101145,
                     yorkcounty_202112141030,
                     yorkcounty_202112141045,
                     yorkcounty_202112141100,
                     yorkcounty_202112141115,
                     yorkcounty_202112141130,
                     yorkcounty_202112141145,
                     yorkcounty_202112141200,
                     yorkcounty_202112141215,
                     	yorkcounty_202112141230,
                     yorkcounty_202112291145,
                     yorkcounty_202112291200,
                     yorkcounty_202112291215,
                     yorkcounty_202112291230,
                     	yorkcounty_202112291245,
                     yorkcounty_202112291300,
                     yorkcounty_202112291315,
                     yorkcounty_202112291330,
                     yorkcounty_202201031600,
                     yorkcounty_202201101000,
                     yorkcounty_202201101015,
                     yorkcounty_202201101030,
                     	yorkcounty_202201101045,
                     yorkcounty_202201101100,
                     	yorkcounty_202201101115,
                     	yorkcounty_202201101130,
                     yorkcounty_202201101145,
                     yorkcounty_202201101200,
                     yorkcounty_202201101215,
                     yorkcounty_202201101230,
                     yorkcounty_202201101245,
                     yorkcounty_202201101300,
                     	yorkcounty_202201101315,
                     	yorkcounty_202201101330,
                     yorkcounty_202201101345,
                     yorkcounty_202201101400,
                     yorkcounty_202201101415,
                     	yorkcounty_202201141200,
                     	yorkcounty_202201191045) %>%
  select(-c(index, TimeServedDays,ReasonCode)) %>% # empty or irrelevant columns
  mutate(personID = sapply(strsplit(GeneralIdentifier, ":", fixed=TRUE), head, 1),
         personID = str_trim(personID, side = "both")) %>%
  mutate(CommitDateTime = parse_date_time(CommitDateTime, '%m/%d/%Y %I:%M:%S %p',
                                          tz = "America/New_York"),
         StartDateTime = parse_date_time(StartDateTime, '%m/%d/%Y %I:%M:%S %p', 
                                         tz = "America/New_York"),
         ProjectedReleaseDateTime = parse_date_time(ProjectedReleaseDateTime, 
                                                    '%m/%d/%Y %I:%M:%S %p', 
                                                    tz = "America/New_York"),
         DateTimeRelease = parse_date_time(DateTimeRelease, 
                                                    '%m/%d/%Y %I:%M:%S %p', 
                                                    tz = "America/New_York"))
  
```


Add personID to historical bookings table.
```{r}
bookings_clean <- bookings %>%
  select(-c(index, TimeServedDays,ReasonCode)) %>% # empty or irrelevant columns
  mutate(personID = sapply(strsplit(GeneralIdentifier, ":", fixed=TRUE), head, 1),
         personID = str_trim(personID, side = "both"))
         
```


```{r}
allbookings <- rbind(bookings_clean, sftp_bookings)
allbookings <- distinct(allbookings, across())
  
```

### Combine MH status & bookings
"The MHSR rating is not assigned to a booking, but rather the patient themselves. So there is no direct relation between a rating and a booking. If a patient comes in on January 5th and gets an A rating, is released on January 8th, they will maintain that rating the next time they are booked in. Unfortunately, due to this, there is no direct correlation between a booking number and a rating, there is only a rating assigned to the patient and the date it was assigned. This is all independent of the booking itself."

* assume the dates of patient rating and booking start/end are always accurate and a person's rating never gets lost
* Interweave & order the dates of bookings & MH ratings per individual.
* If a rating date occurs during a booking (must be between commit & release, inclusive), match it.
* assign each booking after it a rating based on the "statusdate" that is either within it, or the latest prior.
```{r}
# combine all dates in 1 tibble
ratingdates <- inmates %>% arrange(statusdate) %>% 
  select(pouch, statusdate) %>% 
  mutate(pouch = as.character(pouch),
         statusdate = as.character(statusdate))


booking_ranges <- allbookings %>% 
  arrange(CommitDateTime) %>%
  left_join(ratingdates, by = c("personID"="pouch")) %>%
# This would keep only what's in the group-by, an add a list-col
  group_by(personID, GeneralIdentifier, CommitDateTime, DateTimeRelease, ReleaseType, ProjectedReleaseDateTime) %>%
  summarize(statusdates = list(statusdate)) %>%
  ungroup()
  
  #keeps whole table, adds a list-col
  #   rowwise() %>%
  # mutate(statusdates = list(as.Date(statusdate))) %>%
  # ungroup()
  
# This would keep the whole table but add a list col that's a TIBBLE
# nest(statusdates = c(statusdate))
```

#### Helper function, to match the date somepone was assigned a MHSR rating, to the booking it applies to.
Cannot use `case_when` for the `matchdate`. When `in_booking` is empty: "case_when returns a vector with a length equal to the length of the vector you pass in. So nothing in, nothing out."
Cannot use dplyr if_else, it does not evaluate `character(0)` or `logical(0)` well for T/F
```{r}
rating_to_booking <- function(allratings, start=CommitDateTime, end=DateTimeRelease) {  
  end <- if_else(!is.na(end), end, Sys.time()) #if the relase date is null, compare to today's date.
  
  if (!is.na(allratings)) {
    # Check if there's a statusdate during a booking. If so, return it.
  check_in_booking <- between(as.Date(allratings),start,end)
  allratings <-  as.character(allratings)
  
  in_booking <- case_when(
    all(check_in_booking==FALSE) ~ NA_character_,
     TRUE ~ allratings[check_in_booking] %>% as.character()  
  )
    
  # Use the `in_booking` statusdate within the jail booking if exists
  # Or if all statusdates are after the booking ends, there are none
  # Or, get most recent statusdate before the start
  matchdate <- ifelse(length(in_booking) > 0,
                       in_booking, 
                       ifelse(all(as.Date(allratings) < as.Date(end)),
                               allratings[max(rev(order(allratings)))] %>% as.character(), # the max date
                               NA_character_ )
                       )
  return (matchdate)
  }
  
  return (NA_character_) # if no rating, return NA
}
```


<!-- TESTING GROUNDS -->
<!-- ```{r} -->
<!-- # ratingdates = test_ratingdates$statusdate -->
<!-- test =  -->
<!--   booking_ranges2$statusdates[[1]] -->

<!-- start = booking_ranges2$CommitDateTime -->
<!-- end =  booking_ranges2$DateTimeRelease -->

<!-- # if_else(between(testdate, as.Date(start), as.Date(end)), testdate, NA_Date_) -->
<!-- # all(test_ratingdates$statusdate > as.Date(end)) -->
<!-- # between(test_ratingdates$statusdate, as.Date(start), as.Date(end)) -->

<!-- rating_to_booking(allratings = test, -->
<!--                   start = start, -->
<!--                   end = end) -->

<!-- ``` -->

Match MHSR ratings to bookings
```{r}
booking_ranges2 <- booking_ranges %>%
  rowwise() %>%
  mutate(statusdate_match = pmap(list(statusdates[[1]], CommitDateTime, DateTimeRelease),
                                 rating_to_booking)) %>%
  unnest(cols = statusdate_match)
# booking_ranges2$statusdate_match
```

```{r}
alljail_data <- booking_ranges2 %>%
  # join all person names (some won't have statuses)
  left_join(inmates %>% select(pouch, first, last) %>% mutate(pouch = as.character(pouch)),
            by = c("personID"="pouch"),
            na_matches = "never") %>%
  distinct(.) %>% # clear out duplicates, not sure why created
  # join only those statuses that match
  left_join(inmates %>% select(pouch, statusdate, status, id) %>%
              mutate(pouch = as.character(pouch),
         statusdate = as.character(statusdate)), 
            by = c("personID"="pouch", "statusdate_match"="statusdate"),
         na_matches = "never") %>%
  distinct(.) %>% # clear out duplicates, not sure why created
  relocate(statusdates, .after = last_col())
```

### Finish cleaning up bookings table.
* Parse out in their own columns: personID, year booked, timeserved (days)
* Using the *CommitDateTime* and not the StartDateTime to mark the start of the "time served" range.
* If not yet released, calciulate time served until the current date.
```{r}
alljail_data2 <- alljail_data %>%
  mutate(yr_booked_num = year(CommitDateTime),
         yr_booked = as.character(yr_booked_num), #need a text col for listing the "Total" in dropdowns
         ol_timeserveddays = round(if_else(!is.na(DateTimeRelease),
                                           difftime(DateTimeRelease, CommitDateTime, units = "days"),
                                           difftime(Sys.Date(), CommitDateTime, units = "days")),
                                   0),
         statusdates = sapply(statusdates, toString)
         ) %>%
  left_join(releasetype, by="ReleaseType") %>%
  relocate(Description, .after = ReleaseType) %>%
  relocate(c(yr_booked_num, yr_booked, ol_timeserveddays), .after = CommitDateTime) %>%
  relocate(statusdates, .after = last_col())

```

### CALCULATE RECIDIVISM - with 365 day lookback
Data. https://stackoverflow.com/questions/66254603/using-datetime-lookback-periods-to-record-whether-event-occured-in-r
```{r}
# Helper function
previous_booking <- function(person, start, end, data=alljail_data2){
# for each commit, look back 365 days for another commit in that time.
  commits <- data %>% filter(personID==person) %>%
    pull(CommitDateTime)
  
  if_else(commits >= start & commits < end, 1, 0) %>%
    sum() -> flag_count
  
  ifelse(flag_count > 0, 1, 0)
}

alljail_withrecidivism <-  alljail_data2 %>%
  select(personID, CommitDateTime, yr_booked, GeneralIdentifier, status, everything()) %>%
  distinct(personID, CommitDateTime, GeneralIdentifier, status, .keep_all = TRUE) %>%
  arrange(personID, CommitDateTime) %>%
  mutate(
    lookbacktime = (CommitDateTime - years(1)), 
    lookbackflag = pmap_dbl(list(personID, lookbacktime, CommitDateTime), previous_booking)
  )
```


save csv into the `data` folder. 
```{r}
write_csv(alljail_withrecidivism, "data/york_jaildata_combined.csv" )
```


